\documentclass[a4paper]{article}
\title{John bums sheep - an initial investigation}
\author{Fintan S. Nagle}



%BEGIN MATRIX GUMF
% Load TikZ
\usepackage{tikz}
\usetikzlibrary{matrix,decorations.pathreplacing,calc}

% Set various styles for the matrices and braces. It might pay off to fiddle around with the values a little bit
\pgfkeys{tikz/mymatrixenv/.style={decoration=brace,every left delimiter/.style={xshift=3pt},every right delimiter/.style={xshift=-3pt}}}
\pgfkeys{tikz/mymatrix/.style={matrix of math nodes,left delimiter=[,right delimiter={]},inner sep=2pt,column sep=1em,row sep=0.5em,nodes={inner sep=0pt}}}
\pgfkeys{tikz/mymatrixbrace/.style={decorate,thick}}
\newcommand\mymatrixbraceoffseth{0.5em}
\newcommand\mymatrixbraceoffsetv{0.2em}

% Now the commands to produce the braces. (I'll explain below how to use them.)
\newcommand*\mymatrixbraceright[4][m]{
    \draw[mymatrixbrace] ($(#1.north west)!(#1-#3-1.south west)!(#1.south west)-(\mymatrixbraceoffseth,0)$)
        -- node[left=2pt] {#4} 
        ($(#1.north west)!(#1-#2-1.north west)!(#1.south west)-(\mymatrixbraceoffseth,0)$);
}
\newcommand*\mymatrixbraceleft[4][m]{
    \draw[mymatrixbrace] ($(#1.north east)!(#1-#2-1.north east)!(#1.south east)+(\mymatrixbraceoffseth,0)$)
        -- node[right=2pt] {#4} 
        ($(#1.north east)!(#1-#3-1.south east)!(#1.south east)+(\mymatrixbraceoffseth,0)$);
}
\newcommand*\mymatrixbracetop[4][m]{
    \draw[mymatrixbrace] ($(#1.north west)!(#1-1-#2.north west)!(#1.north east)+(0,\mymatrixbraceoffsetv)$)
        -- node[above=2pt] {#4} 
        ($(#1.north west)!(#1-1-#3.north east)!(#1.north east)+(0,\mymatrixbraceoffsetv)$);
}
\newcommand*\mymatrixbracebottom[4][m]{
    \draw[mymatrixbrace] ($(#1.south west)!(#1-1-#3.south east)!(#1.south east)-(0,\mymatrixbraceoffsetv)$)
        -- node[below=2pt] {#4} 
        ($(#1.south west)!(#1-1-#2.south west)!(#1.south east)-(0,\mymatrixbraceoffsetv)$);
}

%END MATRIX GUMF
 
 
 

\input{mylibrary}

\begin{document}
\maketitle


Fire is a complex and dynamic phenomenon. Characterised by rapidly shifting patterns of both first- and second-order motion, it evokes rich visual percepts as well as aesthetic responses. We report an investigation of the visual features useful for discrimination between similar fires, the neurocognitive architecture responsible for comparing them, and the specificity of the neural representations involved.


From an evolutionary point of view, mastery of fire was key to human development. Being able to control fire allowed early humans to cook food, defend themselves from predators and survive in cold, challenging environments. Fire was the first of a long line of technologies which release stored energy from fuel and turn it to human purposes; the earliest archaeological evidence of fire use dates back 1.8 million years (REF), with frequent use found from (FILL) million years onwards (REF). Even before this, hominids regularly encountered flame in the form of bushfires, although these were perceived as a threat, not a controllable, exploitable entity.

The evolving human visual system has therefore been exposed to a large amount of flamelike stimuli in the last 1.8 million years. These stimuli have often appeared in dangerous or life-threatening contexts, either posing a threat or aiding survival. In sufficiently extreme situations, such as extreme cold or heavy predation, those early humans who could successfully control fire had an increased chance of survival.

It is therefore natural to enquire whether the human visual system has become adapted in any way to the perception of flamelike stimuli. Does the visual system employ any specific representations or specialised models when attending to fire, does it use the same general-purpose systems employed when observing a novel moving stimulus?

This question recalls the ongoing debate concerning the specialisation of face perception. We find increased activation of the fusiform face area and inferior temporal sulcus while viewing faces; this can be explained either by innate specialisation or learned proficiency. In the same way, observation of fire may recruit neurons and systems which respond preferentially to, and perform better on, flame stimuli. On the other hand, observing fire may stimulate the same neural populations as observing other moving stimuli.

This report aims to answer two questions. Firstly, what types of visual information are important when observing flames; which details are represented, and which thrown away? This choice defines the upstream visual subsystems (for example, that of motion detection, edge detection or shape detection) which are activated by fire stimuli. Secondly, are any of these systems specific to fire? We do not address the question of whether any specificity is learned or innate.

We begin with a survey of the ecological context in which fire is observed and processed. We then describe several yes/no and 2AFC tasks designed to activate the visual subsystems which process fire and measure task performance under removal of different types of visual information from the stimulus.

\section{Fire in the natural environment}

Here we swiftly review the basic characteristics of fire and the contexts in which early humans interacted with it. It is essential to consider the motivations and tasks which lead the visual system to process and represent fire.

We can define the set of dynamic visual stimuli termed "fire" as the light fields emitted from combusting objects. In terms of luminance, they display rapid variation from gentle glows to dazzling flares, which covers virtually the whole photopic range of the visual system; one can imagine that scotopic vision is not employed except for very dim, distant fires. In terms of size, they may subtend any angle from the whole visual field (for a large, close fire) to a point (for a small, distant fire).

In terms of spatial frequency, fire is characterised by 

FFT stuff here

It is important to consider the context in which flames were observed by early humans. We can imagine two main situations: a) avoiding large natural wildfires, and b) controlling small artificial hearth fires for cooking and defence. Good performance at both tasks is essential for survival, but we are not as interested in visual development supporting task performance in situation a), as it is not specific to humans; all large land animals in an environment at risk of wildfires must possess this skill. We consider only the anthrospecific task of controlling small artificial fires.

Fire is employed in many situations, from defensive to culinary to ritualistic. In order to avoid being overly selective, we do not consider a particular task, but the general context in which they can all be placed: the extraction of useful information (or affordance) from visual stimuli. In order to properly control a fire, the observer must estimate properties such as the temperature of various parts, whether the fire is growing or shrinking, whether it is likely to spark or flare up, and in which direction it will spread. Knowing these details accurately is especially important in the early stages of firelighting, when the flame is small and at risk of extinguishing.

We therefore restrict this work to the study of small, artificial fires which do not pose an imminent threat to survival. 

\section{How is fire represented?}




\section{Are these representations specialised?}



Its presence in the environment of the evolving 


We are thinking on a level between the neural and the cognitive, a level insufficiently detailed to model individual neurons but more specific than descriptions of qualia or percepts. Our main primitive is the concept of the \textit{neural representation}, a description of how a particular idea, metric or concept is coded by neural activity or architecture. It is illustrative to provide some examples.


A place cell, which fires when an animal occupies a particular area of its environment, is a representation of that particular location. Simiarly, a head direction cell represents orientation.

Visual information is encoded in many different representations as it passes from retina to cortex.





\begin{singlespace}
\begin{footnotesize}
\begin{twocolumn}
\bibliographystyle{unsrt}
\bibliography{references}
\end{twocolumn}
\end{footnotesize}
\end{singlespace}
\newpage

\input{appendices}



